---
title: "forecastingNotesAG"
author: "Ali Gharouni"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Update the packages

```{r update packages, echo=FALSE}
# update any installed R packages
update.packages(ask = FALSE, checkBuilt = TRUE)

# packages to install for the course
pkgs <- c("brms", "dplyr", "gratia", "ggplot2", "marginaleffects",
          "tidybayes", "zoo", "viridis", "mvgam")

# install packages
install.packages(pkgs)

# INSTALLING AND CHECKING STAN
install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
library(cmdstanr)
check_cmdstan_toolchain()


install_cmdstan(cores = 2)


```

You should now be able to follow the remaining instructions on the Getting Started with CmdStanR page to ensure that Stan models can successfully compile on your machine. A quick way to check this would be to run this script:


```{r codecheck, echo=FALSE}
library(mvgam)
simdat <- sim_mvgam()
mod <- mvgam(y ~ s(season, bs = 'cc', k = 5) +
               s(time, series, bs = 'fs', k = 8),
             data = simdat$data_train)

```


## Notes during the lectures

** Lec 1 **

- smoothing splines is generally what we will talk about.
- weight on spline basis and penalizing it determines the shape of the smoothers (so estimating the weights is the key) so the theory of GLM to GAM is not a huge jump.
- (temporal) Random effect 

- mgcv is designed with penelized basis
- mvgam is built on mgcv it can handdle the NA in the response variabe but if there is NAs in the covariates, you may need to use mgcv, or impute those

- bam : big addetive model







## Notes on Github branching



To work on a GitHub repository without pushing your changes to the original repository (courseX), you can follow these steps:

1. Fork the Repository
Fork the Repository:
Go to the GitHub page of the repository courseX.
Click the "Fork" button at the top right corner of the page. This will create a copy of the repository under your GitHub account.
2. Clone the Forked Repository
Clone the Forked Repository:
Open your terminal or command prompt.
Clone your forked repository to your local machine using the following command:
sh
Copy code
git clone https://github.com/your-username/courseX.git
Replace your-username with your GitHub username.
3. Set Up Remote Repositories
Set Up Remote Repositories:
Navigate to the repository directory:
sh
Copy code
cd courseX
By default, the origin remote points to your fork. To keep track of the original repository, add it as an upstream remote:
sh
Copy code
git remote add upstream https://github.com/original-owner/courseX.git
Replace original-owner with the username or organization name of the original repository owner.
4. Fetch Updates from the Original Repository
Fetch Updates from the Original Repository:
To keep your fork up-to-date with the original repository, fetch the updates from the upstream remote:
sh
Copy code
git fetch upstream
To merge the updates into your local branch:
sh
Copy code
git checkout main
git merge upstream/main
Replace main with the default branch name if it's different.
5. Work on Your Forked Repository
Work on Your Forked Repository:
You can now work on your forked repository and commit your changes locally.
When you're ready, push your changes to your fork on GitHub:
sh
Copy code
git push origin main
Replace main with the branch you're working on if it's different.
6. Do Not Push to the Original Repository
Do Not Push to the Original Repository:
Since your origin remote points to your fork and not the original repository, you won't accidentally push changes to courseX. Ensure you don't add the original repository as a remote for pushing changes.

